Appendix for: "Bad Forecast, Good Decision

### 2021-02-16 manuscript/appendix.Rmd
:pen:

```{r}
am <- am1$df %>% mutate(belief = am1$posterior$V2)
am %>%  ggplot(aes(time, states[state], col = belief)) + 
  geom_line() + 
  geom_point() +
  scale_colour_gradient(limits = c(0,1), low = pal[1], high = pal[4])
```

@@

```{r plot_sims}
fig2a <- 
  sims %>% 
  group_by(model, time) %>%
  summarise(mean_state = mean(state), sd = sd(state), .groups = "drop") %>%
  filter(time < 25, model != "3") %>%
  ggplot(aes(time, mean_state)) + 
  geom_line(aes(col = model), lwd=1.5, show.legend = FALSE) +
    geom_ribbon(aes(ymin = mean_state - 2*sd, 
                    ymax = mean_state + 2*sd,
                    fill = model),
                alpha = 0.2, show.legend = FALSE) +
  ylab("state") + 
  ggtitle("A") + labs(subtitle = "Ecological")
```

@@

fig2b <- 
npv_df %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time")  + 
  ggtitle("B.") + labs(subtitle = "Economic")

@@

fig3a <- 
  d %>% pivot_longer(names(models), "model") %>%
  ggplot(aes(state, value, col=model, lty=model)) +
  geom_hline(aes(yintercept = 0), lwd=1) + 
  geom_line(lwd=2, show.legend = FALSE) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16)) +
  ylab(bquote(f(x) - x)) + xlab("x")
```

@@

```{r plot_policies}
fig3b <- policies %>%
  ggplot(aes(states, escapement, col=model, lty=model)) + 
  geom_line(lwd=2) + xlab("state")
```

>>>>

```{r}
am <- am1$df %>% mutate(belief = am1$posterior$V2)
am %>%  ggplot(aes(time, states[state], col = belief)) + 
  geom_line() + 
  geom_point() +
  scale_colour_gradient(limits = c(0,1), low = pal[1], high = pal[4]) +
  ylab("stock size")
```

@@

```{r plot_sims}
plot_ecology <- 
  sims %>% 
  group_by(model, time) %>%
  summarise(mean_state = mean(state), sd = sd(state), .groups = "drop") %>%
  filter(time < 25, model != "3") %>%
  ggplot(aes(time, mean_state)) + 
  geom_line(aes(col = model), lwd=1.5, show.legend = FALSE) +
    geom_ribbon(aes(ymin = mean_state - 2*sd, 
                    ymax = mean_state + 2*sd,
                    fill = model),
                alpha = 0.2, show.legend = FALSE) +
  ylab("state") + 
  labs(subtitle = "Ecological")
```

@@

plot_econ <- 
npv_df %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time")  + 
  ggtitle("B.") + labs(subtitle = "Economic")

```
```{r}
# tabular comparisons
npv <- npv_df %>% group_by(model) %>% summarize(npv = max(mean_npv))

am_npv <- sum(am$value * discount ^ am$time)
am_npv / npv[[1,"npv"]]

mean_state <- sims %>% group_by(model) %>% summarize(state = mean(state))
mean(states[am$state]) / mean_state[[1, "state"]]
```

@@

plot_models <- 
  d %>% pivot_longer(names(models), "model") %>%
  ggplot(aes(state, value, col=model, lty=model)) +
  geom_hline(aes(yintercept = 0), lwd=1) + 
  geom_line(lwd=2) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16)) +
  ylab(bquote(f(x) - x)) + xlab("x")
```

@@

```{r plot_policies}
plot_policies <- policies %>%
  ggplot(aes(states, escapement, col=model, lty=model)) + 
  geom_line(lwd=2) + xlab("state")
```

### 2021-02-16 manuscript/appendix_B.Rmd
:pen:

```{r}
names(am1$posterior) <- Ks
belief <- am1$posterior %>%
  mutate(time = 1:dim(am1$posterior)[1]) %>% 
  pivot_longer(as.character(Ks), "K", values_to = "belief") %>%
  mutate(K = as.numeric(K))

belief %>% filter(time %in% seq(1, 11, by = 2)) %>%
  ggplot(aes(K, belief, col=time, group=as.character(time))) +
  geom_line()

belief %>% filter(K < 10) %>%
  ggplot(aes(time, belief, col=K, group=as.character(K))) +
  geom_line()

@@

@@

# Simulations and step-ahead forecasts

We simulate fishing dynamics under the optimal policy for each model, using a simple helper function from the `mdplearning` package. Because growth dynamics are stochastic, we perform 100 simulations of each model from identical starting condition to ensure results are not the result of chance alone.

```{r simulations}
library(mdplearning)
Tmax <- 100
x0 <- which.min(abs(states - 6))
reps <- 100
set.seed(12345)

## Simulate each policy reps times, with `3` as the true model:
simulate_policy <- function(i, policy){
  mdp_planning(transitions[[true_model]], reward, discount,
           policy = policy, x0 = x0, Tmax = Tmax) %>%
    select(value, state_index = state, time, action_index = action)  %>% 
    mutate(state = states[state_index])
}

sims <- 
  map_dfr(names(transitions), 
          function(m){
            policy <- policies %>% filter(model == m) %>% pull(policy)
            map_dfr(1:reps, simulate_policy, policy = policy, .id = "reps")
          },
          .id = "model"
         )
```

```{r stepahead_unfished}
stepahead_unfished <- sims
stepahead_unfished$state_index <- rep(sims$state_index[sims$model == "1"],3)

stepahead_unfished <- stepahead_unfished  %>% 
  filter(model != "3") %>% 
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(expected = transitions[[model]][state_index, , 1]  %*% states,
         var = transitions[[model]][state_index, , 1]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,1]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,1]) > 0.975)) ],
         true = states[next_state ]) 
```

```{r stepahead_fished}
stepahead_fished <- sims %>% 
  filter(model != "3") %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index],
         expected = transitions[[model]][state_index, , action_index]  %*% states,
         var = transitions[[model]][state_index, , action_index]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,action_index]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,action_index]) > 0.975)) ],
         true = states[next_state]) %>%
 select(time, model, true, expected, low, high, var, prob, reps)
```

# Proper scores

```{r proper_scores, message = FALSE}
# Gneiting & Raferty (2007), eq27
scoring_fn <- function(x, mu, sigma){ -(mu - x )^2 / sigma^2  - log(sigma)}

stepahead_unfished <- stepahead_unfished %>%
  mutate(sd = sqrt(var),
         score = scoring_fn(expected, true, sd))

stepahead_fished <- stepahead_fished %>%
  mutate(sd = sqrt(var),
         score = scoring_fn(expected, true, sd))
```

```{r}
predictions <- 
  stepahead_unfished %>% 
  select(time, model, reps, expected, low, high, true, score) %>% 
  mutate(scenario = "A_unfished") %>% 
  bind_rows(stepahead_fished  %>% 
          select(time, model, reps, expected, low, high, true, score) %>% 
          mutate(scenario = "B_fished")
  ) %>% 
  mutate(model = as.character(model))
```

# Plotting

Having computed a single data frame of simulation data under both optimally fished (according to each model) and unfished scenarios for each model, along with the proper scores for the step-ahead forecasts a la @Gneiting2007, we have all the results necessary to generate the figures presented in the main paper.

```{r fig1a}
fig1ab <- predictions %>%
  filter(reps == "2", time < 10) %>%
  ggplot(aes(time, col = model, fill = model)) + 
  geom_point(aes(y = expected)) + 
  geom_errorbar(aes(ymin = low, ymax = high)) +
  geom_point(aes(y = true), pch = "*", size = 12, 
             alpha = 0.6) + 
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "A. Without fishing",
    B_fished = "B. Managed harvest"))) +
  ylab("stock size")

Likewise we can plot the data on proper scores associated with each prediction of each model:

```{r}
fig1cd <- predictions %>%
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 2, show.legend = FALSE) +
  coord_cartesian(xlim = c(-100, 1), ylim = c(0,4000)) +
  xlab("Proper score") +
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "C.",
    B_fished = "D."))) + 
  scale_x_continuous(breaks = c(-100, 0)) +
  theme(axis.text.y = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
        )

## Forecast ecological and economic performance

The ecological performance of the model can easily be visualized by plotting
the results of each simulation.

```{r plot_sims}
fig2a <- 
  sims %>% 
  group_by(model, time) %>%
  summarise(mean_state = mean(state), sd = sd(state), .groups = "drop") %>%
  filter(time < 25, model != "3") %>%
  ggplot(aes(time, mean_state)) + 
  geom_line(aes(col = model), lwd=1.5, show.legend = FALSE) +
    geom_ribbon(aes(ymin = mean_state - 2*sd, 
                    ymax = mean_state + 2*sd,
                    fill = model),
                alpha = 0.2, show.legend = FALSE) +
  ylab("state") + 
  ggtitle("A") + labs(subtitle = "Ecological")
```

To plot the economic value over time, we must sum up the discounted values
at each time step, and then average over replicate simulations of each model:

```{r plot_npv}
##  Net Present Value accumulates over time
npv_df <- sims %>% 
  group_by(model, reps) %>%
  mutate(npv = cumsum(value * discount ^ time)) %>%
  group_by(time, model)  %>% 
  summarise(mean_npv = mean(npv), .groups="drop") %>% 
  arrange(model, time)

optimal <- select(filter(npv_df, model == "3"), time, mean_npv)

fig2b <- 
npv_df %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time")  + 
  ggtitle("B.") + labs(subtitle = "Economic")

Plotting the model functions themselves are shown in Fig 3a:

# References

>>>>

@@

```{r}

am_npv <- sum(am$value * discount ^ am$time)
am_npv / npv[[1,"npv"]]
```

@@
